The Ecology of Cloud Computing

### 2022-04-21 paper.tex
Updated some grammar and added tables

In regards to the climate emergency, 'the cloud' is a rather apt metaphor for network computing; we are yet to learn whether it will provide us shade from a burning issue, or stand as an omen for a terrible storm. It has the potential to reduce the climate impact of several industries, but if not managed carefully it could cause our global emissions to skyrocket. This report will survey the current climate impact of scientific computing with case studies in particular pertinent areas and parts of scientific computing workflows, taking a particular focus on cloud computing and how we can take a variety of approaches to mitigate the extra overhead introduced by cloud facilities, finishing with a discussion on some present and future ideas on how to reduce our climate impact. \newline

This is a topic I have chosen due to my interest in sustainability, and noting that there is thus far very little research and discussion into the impact of computing on the environment; it is often seen as something which just happens on the screen and has no real existence outside the machine's case. Leading the current research into the field is Lo{\"\i}c Lannelongue, from whose work much of my understanding and research into the topic germinated; his contribution to a digestible and layman-friendly yet uncompromising introduction to the field is invaluable.

\begin{figure}[!ht]
\begin{center}
    \includegraphics[width=.6\textwidth]{Diagrams/GLOBE---1850-2021-MO-withlabels.png}
        \caption{A graphic of global average temperature per year from 1850 to 2021 (I believe this image should be somewhere in every paper on anthropogenic climate change). \citep{hawkins2021show}}
\end{center}
\end{figure}

\subsection{What is cloud computing?}
Cloud computing is simply the ability to use computing resources (like storage or processing space) without having to directly own or manage them. For example, if we wanted to have our own website, we could buy a rack of servers ourselves, plug them into the wall, set them up, and voila: we have a website. But what happens when our website gets popular? What happens if one million people all want to use it at one time? We could keep scaling up, fill our kitchen with server racks, start buying air-conditioning for them, and deal with the rather large electricity bill - an approach that many now-huge websites once had to contend with, but eventually you run out of space and time to look after it all without significant investment. \newline

@@

\subsection{Some intuition for carbon use}
Most estimates of emissions are in the units CO2e, or "Carbon dioxide equivalent". This is to provide a standardised unit for emissions from different gases - for example, the climate impact of 1kg of methane is the same as the impact of 25kg of CO2 \citep{forster2007changes}, so if a process emits 3kg of methane and 4kg of CO2 it has emissions of $(3*25)+4=79$kg CO2e. Still, we generally do not have good intuition when it comes to knowing what 1g CO2e actually \emph{means}. Here are some benchmarks (the first three taken from \citet{lannelongue2021ten}):
\begin{itemize}
\item In a European car, driving one kilometre emits about 175g CO2e.
\item Flying from Paris to London emits around 50kg CO2e.
\item Streaming video for one hour emits around 55g CO2e (although this is hard to keep track of; electricity use from data transmission has halved every 2 years on average \citep{aslan2018electricity}).
\item The entire life-cycle of an iPhone 11 used for 3 years is 72kg CO2e (79\% of which is from manufacturing). \citep{apple2019product}
\item One kWh of electricity from the UK grid emits 233g CO2e. \citep{uk2020greenhouse}
\end{itemize}

@@

\section{Scientific computing and energy}
Scientific computing is, essentially, the process of turning electricity into data. To this end, it can be difficult to reduce the amount of electricity used, as this often involves generating less data. Nonetheless, we can review where current and potential problems lie, and where there is room for improvement; it may be useful to view from an efficiency standpoint, where our "science per unit energy" can be improved.

@@

One final point is to consider the programming language used when it comes to efficiency benefits. \citet{heer2018speed} compares the time taken for various programming languages to estimate $\pi$ via the Leibniz method. He finds that compiled\footnote{'Compiled' means a language where a \emph{compiler} takes your code and turns it into an executable file, which you then run; 'interpreted' means the program is run directly as it goes along, line-by-line by an \emph{interpreter}.} languages such as Go or C++ are the fastest, taking 6-9.5 seconds, whereas some popular interpreted languages such as R or Python3 take 181-243 seconds. Of course, this is not to say that all code should be written in C++ lest we kill the planet. Not everyone knows every programming language, and there are non-speed-related benefits that put one language ahead of the other (like Python's accessibility and large number of addon modules, or R being the \textit{de facto} standard for statistics in the life sciences). Merely, we should more heavily consider it when assessing what language to start a project in. Carbon cost may sway the balance when we think about how many saved processing hours are worth the extra development hours\footnote{In particular, code for an experiment that only runs once may not be worth the extra development time of a more difficult language, but code for a long-lasting piece of sotware may be.}. \newline

\subsection{Data}
Once we have processed our data, what do we do with it? Data storage counts for around 10\% of data center energy usage, for a total of 20TWh per year \citep{shehabi2016united} - nearly 40x the energy used by all personal computers in the UK in a year.  \citep{waters2019energy} Moreover, large amounts of energy is needed to produce all this storage media - the manufacture of a Dell PowerStore9000T hard drive (one of their largest available) produces approximately 2765kg CO2e. \citep{dell2021powerstore} A possible and quite successful solution used for data archival is to move unused data to magnetic tapes; with careful discernment of what data needs to be at one's fingertips versus on a library shelf. Using tape can reduce the space and utilities needed for storage sixfold. \citep{moore2007disk}. \newline

@@

\subsection{Machine learning}
Machine learning is becoming \textit{en vogue} in scientific computing; it is being used for everything from improving early stage drug discovery \citep{imrie2021generating} to solving systems of differential equations. \citep{bhattacharya2020model} In essence, machine learning consists of two stages. The first is a 'training' stage, where a large dataset is processed by an algorithm in a machine; the second is a 'application' stage where the machine then uses what it has learned to make estimations about new data. For example, if I were training a machine to read handwritten letters (see \citet{deng2012mnist}), the 'training' dataset would be a big set of handwritten letters along with what they represent (telling the machine 'these 50 handwritten digits are all the letter 'a' etc.), and the machine would use an algorithm to figure out similarities and differences between letters. The 'application' stage would then be giving it a handwritten letter it hasn't seen before and it figuring out what letter it is from the data it gathered during the 'training' stage. This means machine learning can do a lot of busywork that we previously had to employ humans to do, like picking out good candidates for new drugs or typing up handwritten letters; essentially, it could be seen as a way of 'automating intuition'. \newline

This learning stage can take a lot of computation - the MNIST database from \citet{deng2012mnist} has 60,000 training letters. The Enron corpus (\citet[see][]{klimt2004enron}), a dataset of emails used for English language machine learning, contains 700,000 emails; each of these needs an algorithm to be run on it individually. The total process can take days, and for advanced applications can use  huge amounts of electricity. \citet{strubell2019energy} found that training one model using Google's state-of-the-art BART machine learning framework takes 79 hours, in the process using 1507kWh\footnote{This calculation accounted for the cost of cooling and maintaining the data center during this time, not just the actual processing.} (approximately the energy used by a UK household in six months \citep{waters2019energy}) and 326kg CO2e (!). Smaller frameworks, like the T2T framework, used around 27kWh and 5kg CO2e (which is still a non-negligible amount). \newline

While machine learning has been an amazing development in several fields, it runs the risk of overuse; some software concepts can almost become fashionable, with not enough consideration put into them. Note that the statistics in the previous paragraph were for \emph{training} a model - that is, we generated over 5kg CO2e and \emph{have not even solved any problems yet}. Compared to a non-machine learning approach, it is entirely possible that this lump sum of carbon right at the start will put its footprint far ahead of a non-learning-based software for essentially its entire lifetime - care must be taken that machine learning is the right tool for the job.

@@

This could be supplemented with an example.\footnote{NB: \emph{just} an example. Various other practical concerns vastly complicate this in real life; this simply illustrates the interaction of scientific equipment and computing.} The instruments at STFC's ISIS facility consume around 1.1MW.\citep{findlay2021practical} There are 38 instruments, so we take a rough average of 20KW per instrument. The Intel Xeon X3430 server, which is the highest-power-drawing core listed on green-algorithms.org \citep{lannelongue2021green}, draws 23.8W per core when active; for high performance computing, one may want to use 8 or 16 cores - we will go with 16 and take 380W power draw for the whole server. As a further assumption (we are going for a bad case here), we will say this server is in a data centre, so add the overhead of the datacentre (per server), which will make it 532W. This gives us an approximately 1:38 ratio of instrument power use versus computing power use. Now, here we have a fascinating figure indeed; if we run an algorithm for less than 38 hours and as a result need one fewer hour of instrument use, we have saved energy (and thus carbon!). \newline

This statistic can be a useful tool to consider priorities in where this simulation step can be leveraged. For example, if a department is trying to decide which simulation tool to put funding into, one with a 1:40 ratio may be more attractive than one with a 1:5 ratio. Although, one disadvantage is that this statistic means nothing concrete on an objective level. For a molecular analysis experiment, 1:40 may be very attractive (as one could feasibly reap those time-saving benefits), whereas for a bioinformatics study the complexity of biological processes like protein interactions would make it a struggle to do a significant simulation in that time. \newline

It is important to remember that a lot of scientific equipment (such as the ISIS facility) will likely never be green until the entire national grid is green; they simply use too much power. We can try to reduce it where we can (via methods like this), but to disparage the electricity use of these facilities in general belies the huge environmental benefits created through the scientific studies they make possible; the research into things like superconductors \citep{shang2018time} or creating sustainable synthetic versions of natural resources that is only possible through huge facilities like this.

@@

\subsection{Data centres}
A key factor in considering cloud computing is the idea of 'overhead' - around 40\% of the energy use of a datacenter is purely the energy used to cool it so the servers do not overheat,  \citep{lannelongue2021ten} which would not be necessary in a local server (which is small enough to be cooled effectively by similar techniques as those used to cool human beings). However, as datacentres are geared purely towards the processing and storage of data, this can be done more efficiently. In fact, a 2013 case study (\citet{masanet2013energy}) suggested that by moving common applications used in a company from a local server to the cloud, they could reduce emissions by 87\%\footnote{Not to mention the further reduction in emissions if employees are then able to work from home; if everything is on the cloud anyway, then why would they need to burn petrol to commute?}. Naturally, one large cloud centre can utilise efficiency 'economies of scale' better than dozens of local, small ones. Over time, this has certainly shown; research by Google found that while data centre use increased by 550\% between 2010 and 2018, the amount of energy consumed by datacentres only increased by 6\%, meaning the proportion of global energy used by datacentres did not increase despite their much greater usage. \citep{masanet2020recalibrating}\newline

@@

The PUE of a data centre can be reduced through various considerations; the first is how to manage heat emissions. Cooling the servers using water cooling rather than air conditioning (in fact, air conditioning has recently come under fire for creating a vicious cycle as the energy from air conditioning heats the planet, which means more air conditioning is required, and so on \citep{ospina2018cooling}) can be much more efficient. Design can be used to further improve this; a common way of cooling data centres is by cooling alternating rows of servers while letting the odd rows heat up, which creates a convection current around the whole room that keeps everything cool. The water cooling itself can also be made more effective using ground (passive) cooling - by piping the water underground before piping it back into the data centre, we can take advantage of the effectively unlimited cold under the earth. \newline

@@

where $SC$ is the relative Sneakernet carbon (how much carbon emitted by Sneakernet relative to that emitted by internet transfer), $E_{vehicle}$ is the emissions of the vehicle (in g CO2e/km), $d$ the distance of transfer (in km), $E_{internet}$ the emissions of data transmission (in g CO2e/GB) and $s$ the size of the data being transferred (in GB). A 2016 study estimated the energy cost of data transfer as 0.023kWh per gigabyte. \citep{aslan2018electricity} In the United Kingdom, this would emit 5g CO2e/GB. We will take our vehicle as a Mercedes Sprinter (the van used by Royal Mail), which emits approximately 192.6g CO2e/km. \citep{mercedes2019sprinter} We will say our facility is in central London, and our data centre at the Cyxtera Data Centre Campus, a popular data centre campus in Slough - a distance of around 35km. This creates a linear plot from which we can see Sneakernet's carbon emissions break even to the Internet's at a data size of 1348.2GB; that is just one or two regular-sized hard drives! This is even better with electric vehicles; Royal Mail plans to switch to the electric Arrival van, which at full capacity stores 139kWh with a range of 340km \citep{hubbard2021new}. If charged using fossil-fuel generated electricity (the worst-case scenario) this emits 94g CO2e/km, meaning we 'break even' on SneakerNet at 658GB, which one could probably fit on a single laptop. \newline

@@

Another area of interest is quantum computing - this uses concepts from quantum physics to redo computing in a more efficient way 'from scratch'. Quantum computing involves the use of 'qubits', which take advantage of the uncertainty present in quantum-mechanical information to have more states than the 0 or 1 of a classical computing bit, and leverage this for incredibly fast computation. \citep{national2019quantum} The difference that this will make to the environment is completely unimaginable; it is estimated to be capable of quickly working through algorithms that normal computers could not do between now and the end of the universe, which completely changes the lens through which we see computing efficiency. \citep{gyongyosi2019survey} However, we cannot tell if the infrastructure required will balance it out in early days; as it stands, quantum computers must be supercooled to work, which uses a massive amount of cooling energy even on a very small scale. In fact, a sub-field of research into quantum computing is whether it is possible to make large quantum computers without the excessive heat causing the quantum components to fail. \citep{ikonen2017energy}\newline

@@

There is a default set of questions that we ask when we carry out a project. We ask 'is this going to be useful?', 'is this user-friendly?', or 'is my new building accessible to disabled people?'. Perhaps it is time to add a new question; "does my project work for the environment, or against it?" Can we at least work in harmony with environmental goals, or even better, contribute to them? You don't even have to care about the environment to do this - more eco-friendly businesses have better job satisfaction with less burnout and turnover \citep{yu2020hotels} \citep{reverte2016influence}, better perception from consumers \citep{miles1993eco}, and have an easier time recruiting new and better talent \citep{aranganathan2018green} - the latter two growing more and more important, as younger, more environmentally-conscious individuals begin to join the workforce (and consumer perception is, naturally, potential employee perception).\newline

>>>>

In regards to the climate emergency, 'the cloud' is a rather apt metaphor for network computing; we are yet to learn whether it will provide us shade from a burning issue, or stand as an omen for a terrible storm. It has the potential to reduce the climate impact of several industries, but if not managed carefully it could cause our global emissions to skyrocket. This report will survey the current climate impact of scientific computing with case studies in pertinent scientific computing concepts. We will then take a particular focus on cloud computing, and how we can take a variety of approaches to mitigate the extra overhead introduced by cloud facilities. We will finish with a discussion on some present and future ideas on how to reduce our climate impact. \newline

I have chosen this topic due to my interest in sustainability, and noted that there is thus far very little research and discussion into the impact of computing on the environment; it is often seen as something which just happens on the screen and has no real existence outside the machine's case. Leading the current research into the field is Lo{\"\i}c Lannelongue, from whose work much of my understanding and research into the topic germinated; his contribution to a digestible and layman-friendly yet uncompromising introduction to the field is invaluable. The importance of this discussion does not need to be explained with its importance today; instead, I refer to \autoref{stripes} as a 'graphical abstract' on the field.

\begin{figure}[!ht]
\begin{center}
    \includegraphics[width=.6\textwidth]{Diagrams/GLOBE---1850-2021-MO-withlabels.png}
        \caption{A graphic of global average temperature per year from 1850 to 2021 (I believe this image should be somewhere in every paper on anthropogenic climate change). \citep{hawkins2021show}}
	\label{stripes}
\end{center}
\end{figure}

\subsection{What is cloud computing?}
Cloud computing is simply the ability to use computing resources (like storage or processing space) without having to directly own or manage them. For example, if we wanted to have our own website, we could buy a rack of servers ourselves, plug them into the wall, set them up, and voila: we have a website. But what happens when our website gets popular? What happens if one million people all want to use it at one time? We could keep scaling up, fill a room with server racks, start buying air-conditioning for them, and deal with the rather large electricity bill - an approach that many now-huge websites once had to contend with, but eventually you run out of space and time to look after it all without significant investment. \newline

@@

\subsection{Some intuition for carbon use}
Most estimates of emissions are in the units CO2e, or "Carbon dioxide equivalent". This is to provide a standardised unit for emissions from different gases - for example, the climate impact of 1kg of methane is the same as the impact of 25kg of CO2 \citep{forster2007changes}, so if a process emits 3kg of methane and 4kg of CO2 it has emissions of $(3*25)+4=79$kg CO2e. Still, we generally do not have good intuition when it comes to knowing what 1g CO2e actually \emph{means}. Here are some benchmarks (the first three taken from \citet{lannelongue2021ten}):

\begin{figure}[!ht]
\begin{center}
\begin{tabular}{l l l}
  Activity & CO2e & Notes \\
  \hline\hline
  Driving 1km in a European car & 175g & \\
  Flying from Paris to London & 50kg & \\
  Streaming video for one hour & 55g & $\dag_1$ \\
  The entire life-cycle of an iPhone 11 used for 3 years & 72kg \citep{apple2019product} & $\dag_2$ \\
  One kWh of electricity from the UK grid & 233g \citep{uk2020greenhouse} & \\
\end{tabular}
\caption{The carbon emissions from various electricity-consuming processes.\\
 $\dag_1$: This is hard to keep track of; electricity use from data transmission has halved every 2 years on average \citep{aslan2018electricity}. \\
$\dag_2$: 79\% of this is from manufacturing.}
\end{center}
\end{figure}

@@

\section{Scientific computing and energy}
Scientific computing heavily involves the processing of data collected by scientific instruments. To this end, we are essentially turning electricity into data; this can make it hard to reduce our electricity use, as this would involve generating less data. Nonetheless, we can review where current and potential problems lie, and where there is room for improvement; it may be useful to view from an efficiency standpoint, where our "science per unit energy" can be improved.

@@

One final point is to consider the programming language used when it comes to efficiency benefits. \citet{pereira2017energy} compares the energy consumed by various programming languages running a corpus of (mostly bioinformatic) algorithms. It finds that compiled\footnote{'Compiled' means a language where a \emph{compiler} takes your code and turns it into an executable file, which you then run; 'interpreted' means the program is run directly as it goes along, line-by-line by an \emph{interpreter}.} languages such as C or Rust are lowest energy consumers, whereas some popular interpreted languages such as Python3 or Lua used over 75 times as much energy. Of course, this is not to say that all code should be written in C lest we kill the planet. Not everyone knows every programming language, and there are non-speed-related benefits that put one language ahead of the other (like Python's accessibility and large number of addon modules, or the relatively slow statistics-focused language R being the \textit{de facto} standard for statistics in the life sciences). Merely, we should more heavily consider it when assessing what language to start a project in. Carbon cost may sway the balance when we think about how many saved processing hours are worth the extra development hours\footnote{In particular, code for an experiment that only runs once may not be worth the extra development time of a more difficult language, but code for a long-lasting piece of sotware may be.}. \newline

\subsection{Data}
Once we have processed our data, what do we do with it? Data storage counts for around 10\% of data center energy usage, for a total of 20TWh per year \citep{shehabi2016united} - nearly 40x the energy used by all personal computers in the UK in a year.  \citep{waters2019energy} Moreover, large amounts of energy is needed to produce all this storage media - the manufacture of a Dell PowerStore9000T hard drive (one of their largest available) produces approximately 2765kg CO2e. \citep{dell2021powerstore} A common solution used for data archival is to move unused data to magnetic tapes; with careful discernment of what data needs to be at one's fingertips versus on a library shelf. Using tape can reduce the space and utilities needed for storage sixfold. \citep{moore2007disk}. \newline

@@

\subsection{Machine learning}
Machine learning is becoming \textit{en vogue} in scientific computing; it is being used for everything from improving early stage drug discovery \citep{imrie2021generating} to solving systems of differential equations. \citep{bhattacharya2020model} In essence, machine learning consists of two stages. The first is a 'training' stage, where a large dataset is processed by an algorithm in a machine; the second is a 'application' stage where the machine then uses what it has learned to make estimations about new data. For example, if I were training a machine to read handwritten letters (see \citet{deng2012mnist}), the 'training' dataset would be a big set of handwritten letters along with what they represent (telling the machine "these 50 handwritten digits are all the letter 'a'" etc.), and the machine would use an algorithm to figure out similarities and differences between letters. The 'application' stage would then be giving it a handwritten letter it hasn't seen before and it figuring out what letter it is from the data it gathered during the 'training' stage. This means machine learning can do a lot of busywork that we previously had to employ humans to do, like picking out good candidates for new drugs or typing up handwritten letters; essentially, it could be seen as a way of 'automating intuition'. \newline

This learning stage can take a lot of computation - the MNIST database from \citet{deng2012mnist} has 60,000 training letters. The Enron corpus (\citet[see][]{klimt2004enron}), a dataset of emails used for English language machine learning, contains 700,000 emails; each of these needs an algorithm to be run on it individually. The total process can take days, and for advanced applications can use  huge amounts of electricity. \citet{strubell2019energy} found that training one model using Google's state-of-the-art BERT machine learning framework takes 79 hours, in the process using 1507kWh\footnote{This calculation accounted for the cost of cooling and maintaining the data center during this time, not just the actual processing.} (approximately the energy used by a UK household in six months \citep{waters2019energy}) and 652.3kg CO2e (!). Smaller frameworks, like the T2T framework, used around 27kWh and 5kg CO2e (which is still a non-negligible amount).\newline

While machine learning has been an amazing development in several fields, it runs the risk of overuse; some software concepts can almost become fashionable, with not enough consideration put into them. Note that the statistics in the previous paragraph were for \emph{training} a model - that is, we generated over 5kg CO2e and \emph{have not even solved any problems yet}. Compared to a non-machine learning approach, it is entirely possible that this lump sum of carbon right at the start will put its footprint far ahead of a non-learning-based software for essentially its entire lifetime - care must be taken that machine learning is the right tool for the job. At worst, machine learning can be a carbon-intensive way to create statistical flaws and confounding bias. \citep{mania2019model}

\begin{figure}[!ht]
\begin{center}
\begin{tabular}{l l l l}
  Model & Power (W) & Hours to train & CO2e (kg) \\
  \hline\hline
  $Transformer_{base}$ & 1415.78 & 12 & 11.8 \\
  $Transformer_{big}$ & 1515.43 & 84 & 87.1 \\
  $ELMo$ & 517.66 & 336 & 118.8 \\
  $BERT_{base}$ & 12041.51 & 79 & 652.3
\end{tabular}
\caption{The power draw and CO2 emissions produced by various popular machine learning frameworks, taken from \citet{strubell2019energy}. Note that there is not an exact relationship between time taken and carbon usage, as these frameworks were trained on different-powered hardware to each other.}
\end{center}
\end{figure}

@@

A motivating example\footnote{NB: \emph{just} an example. Various other practical concerns vastly complicate this in real life; this simply illustrates the interaction of scientific equipment and computing.}: the instruments at STFC's ISIS facility consume around 1.1MW.\citep{findlay2021practical} There are 38 instruments, so we take a rough average of 20KW per instrument. The Intel Xeon X3430 server, which is the highest-power-drawing core listed on green-algorithms.org \citep{lannelongue2021green}, draws 23.8W per core when active; for high performance computing, one may want to use 8 or 16 cores - we will go with 16 and take 380W power draw for the whole server. As a further assumption (we are going for a worst-case scenario), we will say this server is in a data centre, so add the overhead of the datacentre (per server), which will make it 532W. This gives us an approximately 1:38 ratio of instrument power use versus computing power use. Now, here we have a fascinating figure indeed; if we run an algorithm for less than 38 hours and as a result need one fewer hour of instrument use, we have saved energy (and thus carbon!). \newline

This statistic can be a useful tool to consider priorities in where this simulation step can be leveraged. For example, if a department is trying to decide which instrument gets simulation funding, one with a 1:40 ratio may be more attractive than one with a 1:5 ratio. Although, one disadvantage is that this statistic means nothing concrete on an objective level. For a molecular analysis experiment, 1:40 may be very attractive (as one could feasibly reap those time-saving benefits), whereas for a bioinformatics study the complexity of biological processes like protein interactions would make it a struggle to do a significant simulation in that time. \newline

It is important to remember that a lot of scientific equipment (such as the ISIS facility) will likely never be green until the entire national grid is green; they simply use too much power. We can try to reduce it where we can (via methods like this), but to disparage the electricity use of these facilities in general belies the huge environmental benefits created through the scientific studies they make possible. Much research into things like superconductors \citep{shang2018time} or creating sustainable synthetic versions of natural resources \citep{griffiths2010pgse} is only possible through huge facilities like this.

@@

\subsection{Data centres}
A key factor in considering cloud computing is the idea of 'overhead' - around 40\% of the energy use of a datacenter is purely the energy used to cool it so the servers do not overheat,  \citep{lannelongue2021ten} which would not be necessary in a local server (one in a building which is small enough to be cooled effectively by similar techniques as those used to cool human beings). However, as datacentres are geared purely towards the processing and storage of data, this can be done more efficiently. In fact, a 2013 case study (\citet{masanet2013energy}) suggested that by moving common applications used in a company from a local server to the cloud, they could reduce emissions by 87\%\footnote{Not to mention the further reduction in emissions if employees are then able to work from home; if everything is on the cloud anyway, then why would they need to burn petrol to commute?}. Naturally, one large cloud centre can utilise efficiency 'economies of scale' better than dozens of local, small ones. Over time, this has certainly shown; research by Google found that while data centre use increased by 550\% between 2010 and 2018, the amount of energy consumed by datacentres only increased by 6\%, meaning the proportion of global energy used by datacentres did not increase despite much greater use. \citep{masanet2020recalibrating}\newline

@@

The PUE of a data centre can be reduced through various considerations; the first is managing heat emissions. Cooling the servers using water cooling rather than air conditioning (in fact, air conditioning has recently come under fire for creating a vicious cycle as the energy from air conditioning heats the planet, which means more air conditioning is required, and so on \citep{ospina2018cooling}) can be much more efficient. \citep{zhang2022cooling} Design can be used to further improve this; a common way of cooling data centres is by cooling alternating rows of servers while letting the odd rows heat up, which creates a convection current around the whole room that keeps everything cool. The water cooling itself can also be made more effective using ground (passive) cooling - by piping the water underground before piping it back into the data centre, we can take advantage of the effectively unlimited cold under the earth. \newline

@@

where $SC$ is the relative Sneakernet carbon (how much carbon emitted by Sneakernet relative to that emitted by internet transfer), $E_{vehicle}$ is the emissions of the vehicle (in g CO2e/km), $d$ the distance of transfer (in km), $E_{internet}$ the emissions of data transmission (in g CO2e/GB) and $s$ the size of the data being transferred (in GB). A 2016 study estimated the energy cost of data transfer as 0.023kWh per gigabyte. \citep{aslan2018electricity} In the United Kingdom, this would emit 5g CO2e/GB. We will take our vehicle as a Mercedes Sprinter (the van used by Royal Mail), which emits approximately 192.6g CO2e/km. \citep{mercedes2019sprinter} We will say our facility is in central London, and our data centre at the Cyxtera Data Centre Campus, a popular data centre campus in Slough - a distance of around 35km. \newline

This creates a linear plot from which we can see Sneakernet's carbon emissions break even to the Internet's at a data size of 1348.2GB; that is just one or two regular-sized hard drives! This is even better with electric vehicles; Royal Mail plans to switch to the electric Arrival van, which at full capacity stores 139kWh with a range of 340km \citep{hubbard2021new}. If charged using fossil-fuel generated electricity (the worst-case scenario) this emits 94g CO2e/km, meaning we 'break even' on SneakerNet at 658GB, which one could probably fit on a single laptop. \newline

@@

Another area of interest is quantum computing - this uses concepts from quantum physics to redo computing in a more efficient way 'from scratch'. Quantum computing involves the use of 'qubits', which take advantage of the uncertainty present in quantum-mechanical information to have more states than the 0 or 1 of a classical computing bit, and leverage this for incredibly fast computation. \citep{national2019quantum} The difference that this will make to the environment is completely unimaginable; it is estimated to be capable of quickly working through algorithms that normal computers could not do between now and the end of the universe! \citep{gyongyosi2019survey} However, we cannot tell if the infrastructure required will balance it out in early days; as it stands, quantum computers must be supercooled to work, which uses a massive amount of cooling energy even on a very small scale. In fact, a sub-field of research into quantum computing is whether it is possible to make large quantum computers without the excessive heat causing the quantum components to fail. \citep{ikonen2017energy}\newline

@@

There is a default set of questions that we ask when we carry out a project. We ask 'is this going to be useful?', 'is this user-friendly?', or 'is my new building accessible to disabled people?'. Perhaps it is time to add a new question: "does my project work for the environment, or against it?" Can we at least work in harmony with environmental goals, or even better, contribute to them? You don't even have to care about the environment to do this; more eco-friendly businesses have better job satisfaction with less burnout and turnover \citep{yu2020hotels} \citep{reverte2016influence}, better perception from consumers \citep{miles1993eco}, and have an easier time recruiting new and better talent \citep{aranganathan2018green} - the latter two growing more and more important, as younger, more environmentally-conscious individuals begin to join the workforce (and consumer perception is, naturally, potential employee perception).\newline
